---
title: "cleaner GJAM"
author: "Sarah Roberts"
date: "3/30/2019"
output: html_document
---

```{r packages, include=FALSE}
library(gjam)
library(tidyr)#data manipulation
library(dplyr) #data manipulation
library(naniar) #replace with NA
```

#Load data 
```{r load data}
#okay we need to start from scratch because some of the bratt stuff was wonky (there were no species names after the letter M)
#first I am going to load the original spring and fall data
#then make in a workable format 

#load data 
fall_catch <- read.csv("~/Documents/ClimateOceanPlanning/Datasets/NEFSC_Bottom_Trawl/22560_FSCSTables_Fall/22560_UNION_FSCS_SVCAT.csv")
fall_x <- read.csv("~/Documents/ClimateOceanPlanning/Datasets/NEFSC_Bottom_Trawl/22560_FSCSTables_Fall/22560_UNION_FSCS_SVSTA.csv")

spring_catch <- read.csv("~/Documents/ClimateOceanPlanning/Datasets/NEFSC_Bottom_Trawl/22561_FSCSTables_spring/22561_UNION_FSCS_SVCAT.csv")
spring_x <- read.csv("~/Documents/ClimateOceanPlanning/Datasets/NEFSC_Bottom_Trawl/22561_FSCSTables_spring/22561_UNION_FSCS_SVSTA.csv")

```

```{r data manip}
#manipulated the catch data to be in tidy format 
fall_catch <- fall_catch %>% 
 group_by_at(vars(-expcatchnum)) %>% # group by everything other than the value column. 
 mutate(row_id=1:n()) %>% ungroup() %>% # build group index
 spread(key=scientific_name, value=expcatchnum)  # spread

fall_catch$id <- as.numeric(as.character(fall_catch$id))

fall_catch_select <- fall_catch %>% 
  select(-catch_comment) %>% 
  mutate_if(is.factor, as.integer) %>% 
  group_by(id) %>% 
  replace(is.na(.), 0) %>%
 summarise_each(funs(max)) 

fall_catch_select <- fall_catch_select %>% 
  select(-status_code, -svspp, -catchsex, -expcatchwt, -row_id)

#joining the tables 
fall_x$id <- as.numeric(as.character(fall_x$id))
fall <- left_join(fall_catch_select, fall_x, by="id")
fall$season <- "fall"

#doing the same for spring 
spring_catch <- spring_catch %>% 
 group_by_at(vars(-expcatchnum)) %>% # group by everything other than the value column. 
 mutate(row_id=1:n()) %>% ungroup() %>% # build group index
 spread(key=scientific_name, value=expcatchnum)  # spread

spring_catch$id <- as.numeric(as.character(spring_catch$id))

spring_catch_select <- spring_catch %>% 
  select(-catch_comment) %>% 
  mutate_if(is.factor, as.integer) %>% 
  group_by(id) %>% 
  replace(is.na(.), 0) %>%
 summarise_each(funs(max)) 

spring_catch_select <- spring_catch_select %>% 
  select(-status_code, -svspp, -catchsex, -expcatchwt, -row_id)

#joining the tables 
spring_x$id <- as.numeric(as.character(spring_x$id))
spring <- left_join(spring_catch_select, spring_x, by="id")
spring$season <- "spring"

#combining spring and fall 
#first lets select only the overlapping species/columns 
cols <- intersect(colnames(spring), colnames(fall))
total <- rbind(spring[,cols], fall[,cols])

#now need to filter for the data after 1982 (this is when Batt uses it. There was a change in otter doors to allow more pelagics, and more consistent sampling of southern stratum) http://www.ices.dk/sites/pub/CM%20Doccuments/1997/Y/1997_Y33.pdf

total <- total %>% 
  filter(est_year>1981)

total_latlondate <- total %>% 
  select(id, decdeg_beglat, decdeg_beglon, est_year, est_month)

write.csv(total_latlondate, file="latlondate_total.csv")

```
Now I need to go back and make a few files. The first one is the NAO/AMO file which is located in the longlatdat_total xl file. Use the index function to get the NAO/AMO values by year. 

The next step is to go in and copy the long lat file to arcmap and use that and the benthic habitat poly 2 shapefile along with the IDENTITY tool to get all of the information into one. 



```{r Adding external data}
#adding AMO/NAO and habitat information 
climate_vars <- read.csv("~/Documents/ClimateOceanPlanning/R/NEFSC_GJAMM_GIT/NAO_AMO_dat.csv")
hab_vars <- read.csv("~/Documents/ClimateOceanPlanning/R/NEFSC_GJAMM_GIT/habitat_dat.csv")

total <- rbind(spring[,cols], fall[,cols])
total <- total %>% 
  filter(est_year>1981)
#remove id's where NA 
total <- total %>% 
  filter(!is.na(id)) 

total <- total %>% 
  filter(!is.na(beglat)) #idk why there was a spatial NA but got rid of it in the other files and here #note this is really frustrating and can't figure out that the id columns are the same so I might just do it in excel 
#note this id is missing spatial points 
#201204013400080000.00


#bind them all together 
total <- cbind(total, climate_vars, by="id")

total <- cbind(total, hab_vars, by="id", all=T)

#getting rid of some overlapping columns 
total <- total[, !duplicated(colnames(total))]

#replacing spaces with _ in column names 
names(total) <- gsub(" ", "_", names(total))

#trimming so just the columns we want 

xdata <- total %>% 
 select(id, est_month, est_day, est_year,surftemp, surfsalin, bottemp,
     botsalin, AMO_unsm_month, AMO_unsm_year, NAO_month, NAO_year, NAO_JFM, avgdepth,
     LPI, SLOPE, SBF_GRP, SBFDEPTH, SEDIMENT, SBFDESED, BENHAB, decdeg_beglat, decdeg_beglon)

#lets also add a column for sediment type 
xdata$SED_TYPE <- ifelse(xdata$SEDIMENT ==  '0 - 0.03 Silt/Mud', 'Silt/Mud', 
                        ifelse(xdata$SEDIMENT ==  '0.03 - 0.17 Sand', 'Sand',
                               ifelse(xdata$SEDIMENT ==  '0.17 - 0.35 Sand', 'Sand', 
                                      ifelse(xdata$SEDIMENT ==  '0.35 - 0.36 Sand', 'Sand',
                                             ifelse(xdata$SEDIMENT ==  '0.36 - 0.48 Sand', 'Sand',
                                                    ifelse(xdata$SEDIMENT ==  '20.03 - 0.17 Sand',
                                                           'Sand',
                                                    ifelse(xdata$SEDIMENT ==  '0.48+ Coarse Sand to  
                                                           Gravel','Gravel', 
                                                           ifelse(xdata$SEDIMENT ==  '60.48+ Coarse 
                                                                  Sand to Gravel',
                                                                  'Gravel', NA))))))))
xdata$SED_TYPE <- ifelse(xdata$SEDIMENT ==  '0.48+ Coarse Sand to Gravel','Gravel', ifelse(xdata$SEDIMENT ==  '60.48+ Coarse Sand to Gravel', 'Gravel', xdata$SED_TYPE))
#weirdly have to do this in two steps, I think it doesnt like too many if else statements oh well. 

#too many to do like this for ydata so will just select the columns 
ydata <- total[,7:513]
```
#checkpoint 
do we have adataset by ID with all of the x and y data as well as a dataset with just x data and just y data? 


#selecting species 
So I could go a couple of ways here. I could follow Batt 2017 which selected a number of species, or I could go on my own and select species that are present in a lot of the trawls. I am going to do the latter first.

```{r selecting species, include=FALSE}
#splitting into x and ydata 

ycommon <- gjamTrimY(ydata, 100, OTHER=F)$y

```


#running GJAM 
splitting into chunks 

```{r first time chunk}
total_data <- cbind(ycommon,xdata)


#first time chunk 
total <- subset(total_data, est_year < 2000)
ydata <- total[,1:197]
ydata[is.na(ydata)] <- 0
ycommon <- gjamTrimY(ydata, 100, OTHER=F)$y
#need to do this here to see ones that werent caught early and are caught later 

xdata <- total %>% 
 select(id, est_month, est_day, est_year,surftemp, surfsalin, bottemp,
     botsalin, AMO_unsm_month, AMO_unsm_year, NAO_month, NAO_year, NAO_JFM, avgdepth,
     LPI, SLOPE, SBF_GRP, SBFDEPTH, SEDIMENT, SBFDESED, BENHAB, decdeg_beglat, decdeg_beglon,
     SED_TYPE)

S <- ncol(ycommon)
ml <- list(ng = 10000, burnin = 2000, typeNames = 'DA')



xdata <- as.data.frame(xdata)
xdata$GRPSED <- as.factor(as.character(xdata$SEDIMENT))
xdata$SED_TYPE <- as.factor(as.character(xdata$SED_TYPE))
xdata$est_year <- as.factor(xdata$est_year)
xdata$AMO_unsm <- as.numeric(as.character(xdata$AMO_unsm_year))
xdata$avgdepth <- as.numeric(xdata$avgdepth)
xdata$surfsalin <- as.numeric(xdata$botsalin)
xdata$bottemp <- as.numeric(xdata$bottemp)
xdata$NAO <- as.numeric(as.character((xdata$NAO_year)))


out <- gjam(~ avgdepth + bottemp + botsalin + AMO_unsm_year + NAO_year + SED_TYPE, xdata, ycommon, modelList = ml)


save(out, file='out_82_99.Rdata')
load('out_82_99.Rdata')
plotPars <- list(SMALLPLOTS = T, GRIDPLOTS=T)
gjamPlot(out, plotPars=plotPars)

#get informative stuff and save 
DIC_1 <- out$fit$DIC         # DIC
rmspe_1 <- out$fit$rmspeAll    # root mean squared prediction error for all species
rmspe_spec_1 <- out$fit$rmspeBySpec # RMSPE for each species

VIF_1 <- out$inputs$designTable               # VIF for each predictor
beta_1 <- out$parameters$betaMu                # mean beta for each species and predictor
cred_1 <- out$parameters$betaStandXTable # credible interval for each species and predictor
sens_1 <- out$parameters$sensTable             # sensitivity


#lets try and get out only the significant betas 
#need to add a column with the specnames in it because right now its just the row names 
cred_1$names <- rownames(cred_1)

cred_1_sig <- cred_1 %>% 
  filter(sig95 == "*") %>% #only keep sig values
  filter(!grepl('intercept', names))  #remove anything that says intercept in there 

```

 

```{r second time chunk}
#second time chunk 
total <- subset(total_data, est_year > 1999)
ydata <- total[,1:197]
ydata[is.na(ydata)] <- 0
ycommon <- gjamTrimY(ydata, 100, OTHER=F)$y
#need to do this here to see ones that werent caught early and are caught later 

xdata <- total %>% 
 select(id, est_month, est_day, est_year,surftemp, surfsalin, bottemp,
     botsalin, AMO_unsm_month, AMO_unsm_year, NAO_month, NAO_year, NAO_JFM, avgdepth,
     LPI, SLOPE, SBF_GRP, SBFDEPTH, SEDIMENT, SBFDESED, BENHAB, decdeg_beglat, decdeg_beglon,
     SED_TYPE)

S <- ncol(ycommon)
ml <- list(ng = 10000, burnin = 2000, typeNames = 'DA')



xdata <- as.data.frame(xdata)
xdata$GRPSED <- as.factor(as.character(xdata$SEDIMENT))
xdata$SED_TYPE <- as.factor(as.character(xdata$SED_TYPE))
xdata$est_year <- as.factor(xdata$est_year)
xdata$AMO_unsm <- as.numeric(as.character(xdata$AMO_unsm_year))
xdata$avgdepth <- as.numeric(xdata$avgdepth)
xdata$surfsalin <- as.numeric(xdata$botsalin)
xdata$bottemp <- as.numeric(xdata$bottemp)
xdata$NAO <- as.numeric(as.character((xdata$NAO_year)))


out <- gjam(~ avgdepth + bottemp + botsalin + AMO_unsm_year + NAO_year + SED_TYPE, xdata, ycommon, modelList = ml)


save(out, file='out_99_15.Rdata')
load('out_99_15.Rdata')
plotPars <- list(SMALLPLOTS = T, GRIDPLOTS=T)
gjamPlot(out, plotPars=plotPars)

#getting informative stuff 
out$parameters$fSe
out$fit$DIC
gp <- gjamPlot(output = out, plotPars = plotPars)
gp$clusterIndex
#There are three versions, data, E, and R that correspond to plots that come out. E = environmental response, R = residual covariance

DIC_2 <- out$fit$DIC         # DIC
rmspe_2 <- out$fit$rmspeAll    # root mean squared prediction error for all species
rmspe_spec_2 <- out$fit$rmspeBySpec # RMSPE for each species

VIF_2 <- out$inputs$designTable               # VIF for each predictor
beta_2 <- out$parameters$betaMu                # mean beta for each species and predictor
cred_2 <- out$parameters$betaStandXTable # credible interval for each species and predictor
sens_2 <- out$parameters$sensTable             # sensitivity

#head(out$prediction$presence)        # predicted occurance for each species


#lets try and get out only the significant betas 
#need to add a column with the specnames in it because right now its just the row names 
cred_2$names <- rownames(cred_2)

cred_2_sig <- cred_2 %>% 
  filter(sig95 == "*") %>% 
  filter(!grepl('intercept', names))  #remove anything that says intercept in there 
```


```{r total data}
#all data
total <- total_data
ydata <- total[,1:197]
ydata[is.na(ydata)] <- 0
ycommon <- gjamTrimY(ydata, 100, OTHER=F)$y
#need to do this here to see ones that werent caught early and are caught later 

xdata <- total %>% 
 select(id, est_month, est_day, est_year,surftemp, surfsalin, bottemp,
     botsalin, AMO_unsm_month, AMO_unsm_year, NAO_month, NAO_year, NAO_JFM, avgdepth,
     LPI, SLOPE, SBF_GRP, SBFDEPTH, SEDIMENT, SBFDESED, BENHAB, decdeg_beglat, decdeg_beglon,
     SED_TYPE)

S <- ncol(ycommon)
ml <- list(ng = 10000, burnin = 2000, typeNames = 'DA')



xdata <- as.data.frame(xdata)
xdata$GRPSED <- as.factor(as.character(xdata$SEDIMENT))
xdata$SED_TYPE <- as.factor(as.character(xdata$SED_TYPE))
xdata$est_year <- as.factor(xdata$est_year)
xdata$AMO_unsm <- as.numeric(as.character(xdata$AMO_unsm_year))
xdata$avgdepth <- as.numeric(xdata$avgdepth)
xdata$surfsalin <- as.numeric(xdata$botsalin)
xdata$bottemp <- as.numeric(xdata$bottemp)
xdata$NAO <- as.numeric(as.character((xdata$NAO_year)))


out <- gjam(~ avgdepth + bottemp + botsalin + AMO_unsm_year + NAO_year + SED_TYPE, xdata, ycommon, modelList = ml)


save(out, file='out_82_15.Rdata')
plotPars <- list(SMALLPLOTS = T, GRIDPLOTS=T)
gjamPlot(out, plotPars=plotPars)

DIC_tot <- out$fit$DIC         # DIC
rmspe_tot <- out$fit$rmspeAll    # root mean squared prediction error for all species
rmspe_spec_tot <- out$fit$rmspeBySpec # RMSPE for each species

VIF_tot <- out$inputs$designTable               # VIF for each predictor
beta_tot <- out$parameters$betaMu                # mean beta for each species and predictor
cred_tot <- out$parameters$betaStandXTable # credible interval for each species and predictor
sens_tot <- out$parameters$sensTable             # sensitivity

#head(out$prediction$presence)        # predicted occurance for each species


#lets try and get out only the significant betas 
#need to add a column with the specnames in it because right now its just the row names 
cred_tot$names <- rownames(cred_tot)

cred_tot_sig <- cred_tot %>% 
  filter(sig95 == "*") %>% 
  filter(!grepl('intercept', names))  #remove anything that says intercept in there 
```

#checkpoint 
make sure you have the gjams working and all of the significant betas saved into a file 

#shifting env 

```{r matching and shifts}
#find matching cases 
match <- charmatch(cred_1_sig$names, cred_2_sig$names)
cred_1_sig <- cbind(cred_1_sig, match)

#adding columns for descriptions 
cred_1_sig$var <- ifelse(grepl("bottemp", cred_1_sig$names), "temp",
                         ifelse(grepl("botsalin", cred_1_sig$names), "sal", 
                                ifelse(grepl("avgdepth", cred_1_sig$names), "depth",
                                       ifelse(grepl("SEDTYPESilt/Mud", cred_1_sig$names), "silt/mud",
                                               ifelse(grepl("SEDTYPESand", cred_1_sig$names), "sand",
                                                      ifelse(grepl("AMOunsmyear", cred_1_sig$names), "AMO",
                                                             ifelse(grepl("NAOyear", cred_1_sig$names),
                                                                    "NAO", NA)))))))

cred_1_sig$temp <- ifelse(cred_1_sig$Estimate > 1 & cred_1_sig$var == "temp", "warm", 
                           ifelse(cred_1_sig$Estimate < 1 & cred_1_sig$var == "temp", "cool", NA))  

cred_1_sig$sal <- ifelse(cred_1_sig$Estimate > 1 & cred_1_sig$var == "sal", "saline", 
                           ifelse(cred_1_sig$Estimate < 1 & cred_1_sig$var == "sal", "nonsaline", NA)) 

cred_1_sig$depth <- ifelse(cred_1_sig$Estimate > 1 & cred_1_sig$var == "depth", "deep", 
                           ifelse(cred_1_sig$Estimate < 1 & cred_1_sig$var == "depth", "shallow", NA)) 

cred_1_sig$siltmud <- ifelse(cred_1_sig$Estimate > 1 & cred_1_sig$var == "silt/mud", "silt/mud", 
                           ifelse(cred_1_sig$Estimate < 1 & cred_1_sig$var == "silt/mud", "nosilt/mud", NA)) 

cred_1_sig$sand <- ifelse(cred_1_sig$Estimate > 1 & cred_1_sig$var == "sand", "sand", 
                           ifelse(cred_1_sig$Estimate < 1 & cred_1_sig$var == "sand", "nosand", NA))

cred_1_sig$AMO <- ifelse(cred_1_sig$Estimate > 1 & cred_1_sig$var == "AMO", "AMO+", 
                           ifelse(cred_1_sig$Estimate < 1 & cred_1_sig$var == "AMO", "AMO-", NA)) 

cred_1_sig$NAO <- ifelse(cred_1_sig$Estimate > 1 & cred_1_sig$var == "NAO", "NAO+", 
                           ifelse(cred_1_sig$Estimate < 1 & cred_1_sig$var == "NAO", "NAO-", NA)) 


#adding columns for descriptions for the second time period
cred_2_sig$var <- ifelse(grepl("bottemp", cred_2_sig$names), "temp",
                         ifelse(grepl("botsalin", cred_2_sig$names), "sal", 
                                ifelse(grepl("avgdepth", cred_2_sig$names), "depth",
                                       ifelse(grepl("SEDTYPESilt/Mud", cred_2_sig$names), "silt/mud",
                                               ifelse(grepl("SEDTYPESand", cred_2_sig$names), "sand",
                                                      ifelse(grepl("AMOunsmyear", cred_2_sig$names), "AMO",
                                                             ifelse(grepl("NAOyear", cred_2_sig$names),
                                                                    "NAO", NA)))))))

cred_2_sig$temp <- ifelse(cred_2_sig$Estimate > 1 & cred_2_sig$var == "temp", "warm", 
                           ifelse(cred_2_sig$Estimate < 1 & cred_2_sig$var == "temp", "cool", NA))  

cred_2_sig$sal <- ifelse(cred_2_sig$Estimate > 1 & cred_2_sig$var == "sal", "saline", 
                           ifelse(cred_2_sig$Estimate < 1 & cred_2_sig$var == "sal", "nonsaline", NA)) 

cred_2_sig$depth <- ifelse(cred_2_sig$Estimate > 1 & cred_2_sig$var == "depth", "deep", 
                           ifelse(cred_2_sig$Estimate < 1 & cred_2_sig$var == "depth", "shallow", NA)) 

cred_2_sig$siltmud <- ifelse(cred_2_sig$Estimate > 1 & cred_2_sig$var == "silt/mud", "silt/mud", 
                           ifelse(cred_2_sig$Estimate < 1 & cred_2_sig$var == "silt/mud", "nosilt/mud", NA)) 

cred_2_sig$sand <- ifelse(cred_2_sig$Estimate > 1 & cred_2_sig$var == "sand", "sand", 
                           ifelse(cred_2_sig$Estimate < 1 & cred_2_sig$var == "sand", "nosand", NA))

cred_2_sig$AMO <- ifelse(cred_2_sig$Estimate > 1 & cred_2_sig$var == "AMO", "AMO+", 
                           ifelse(cred_2_sig$Estimate < 1 & cred_2_sig$var == "AMO", "AMO-", NA)) 

cred_2_sig$NAO <- ifelse(cred_2_sig$Estimate > 1 & cred_2_sig$var == "NAO", "NAO+", 
                           ifelse(cred_2_sig$Estimate < 1 & cred_2_sig$var == "NAO", "NAO-", NA)) 




#lets just compare the matching ones 
cred_1_sig_match <- cred_1_sig %>% 
   filter(!is.na(match)) 

cred_2_sig$match <- rownames(cred_2_sig)

new <- merge(cred_1_sig_match, cred_2_sig, by = "match")
write.csv(new, file="only_matching_sig_betas.csv")
#went into excel and highlighted the changed ones but also made column called change here is that file 
change <- read.csv("~/Documents/ClimateOceanPlanning/R/NEFSC_GJAMM_GIT/only_matching_sig_betas_changeshown.csv")

#lets get the non matching ones to help us get at question 2 below 
cred_1_sig_nomatch <- cred_1_sig %>% 
   filter(is.na(match)) 
write.csv(cred_1_sig_nomatch, file="cred_1_sig_nomatch.csv")

#other way around also  
match2 <- charmatch(cred_2_sig$names, cred_1_sig$names)
cred_2_sig <- cred_2_sig %>% 
    select(-match) 
cred_2_sig <- cbind(cred_2_sig, match2)

cred_2_sig_nomatch <- cred_2_sig %>% 
   filter(is.na(match2)) 
write.csv(cred_2_sig_nomatch, file="cred_2_sig_nomatch.csv")
```
Okay so we could have a few things happening. 
1. species relationship with BT could be positive in the first time period and negative in second (both significant relationships)
  Highlighted yellow in csv file only_matching_sig_betas_changeshown.csv
2. Species could have significant relationship with BT in first time period and that relationship isnt significant in the second. 
  Those cases are in cred_1_sig_no match (sig in first but not second) or cred_2_sig_nomatch (sig in    second but not first). All of thr changes are in the excel file Change_total. 
    in sigs_all sheet I have highlighted all of the species matching both cases 1 and 2 described above. 

I also have to go through and see if species were not caught (or not caught enought to be included in our gjam analysis) or if the relationship with the environment just wasn't significant. doing that now by highlihgting the ones with no significance in first then something significant in second and looking at all of the gjam beta outputs . 

#shifting mean centroids? 
This is to get at if the environment is changing or if species are changing with the environment. 
```{r load pacakges}
install.packages("sp")
library(sp)
install.packages("spatialEco")
library(spatialEco)
install.packages("geosphere")
library(geosphere)
```
this makes the dataset a spatial object
```{r}
total_sp <- total_data
coordinates(total_sp) <- ~ decdeg_beglon + decdeg_beglat


#Lets select out the time period chunks we need. 
total_sp_1 <- subset(total_data, est_year < 2000)
total_sp_2 <- subset(total_data, est_year > 1999)

#now make all of those spatial objects: 

coordinates(total_sp_1) <- ~ decdeg_beglon + decdeg_beglat

coordinates(total_sp_2) <- ~ decdeg_beglon + decdeg_beglat


```

Now we can only do this for species that were duplicated in all of them. I went back to the excel file (Change_total.xls) and got out all of the species that had significant relationships with the ENV and were duplicated in the two time periods (tab is called specnames). 

```{r}
specnames_1<-c("Alosa_aestivalis_(blueback_herring)", "Alosa_pseudoharengus_(alewife)", "Alosa_sapidissima_(American_shad)", "Amblyraja_radiata_(thorny_skate)", "Ammodytes_dubius_(northern_sand_lance)", "Anchoa_hepsetus_(striped_anchovy)", "Anchoa_mitchilli_(bay_anchovy)", "Argentina_silus_(Atlantic_argentine)", "Aspidophoroides_monopterygius_(alligatorfish)", "Atlantopandalus_propinqvus_()", "Bairdiella_chrysoura_(silver_perch)", "Brevoortia_tyrannus_(Atlantic_menhaden)", "Callinectes_sapidus_(blue_crab)", "Cancer_borealis_(Jonah_crab)", "Cancer_irroratus_(Atlantic_rock_crab)", "Caridea", "Centropristis_striata_(black_sea_bass)", "Chaceon_quinquedens_(red_deepsea_crab)", "Chlorophthalmus_agassizi_(shortnose_greeneye)", "Citharichthys_arctifrons_(Gulf_stream_flounder)", "Clupea_harengus_(Atlantic_herring)", "Crangon_septemspinosa_(sevenspine_bay_shrimp)", "Cryptacanthodes_maculatus_(wrymouth)", "Cynoscion_regalis_(weakfish)", "Dasyatis_say_(bluntnose_stingray)", "Decapterus_punctatus_(round_scad)", "Dichelopandalus_leptocerus_(bristled_longbeak)", "Dipturus_laevis_(barndoor_skate)", "Enchelyopus_cimbrius_(fourbeard_rockling)", "Etropus_microstomus_(smallmouth_flounder)", "Etrumeus_teres_(round_herring)", "Gadus_morhua_(Atlantic_cod)", "Glyptocephalus_cynoglossus_(witch_flounder)", "Helicolenus_dactylopterus_(blackbelly_rosefish)", "Hemitripterus_americanus_(sea_raven)", "Hippoglossina_oblonga_(fourspot_flounder)", "Hippoglossoides_platessoides_(American_plaice)", "Homarus_americanus_(American_lobster)", "Illex_illecebrosus_(northern_shortfin_squid)", "Lagodon_rhomboides_(pinfish)", "Larimus_fasciatus_(banded_drum)", "Leiostomus_xanthurus_(spot)", "Lepophidium_profundorum_(fawn_cusk-eel)", "Leptoclinus_maculatus_(daubed_shanny)", "Leucoraja_erinacea_(little_skate)", "Leucoraja_garmani_(rosette_skate)", "Leucoraja_ocellata_(winter_skate)", "Limanda_ferruginea_(yellowtail_flounder)", "Limulus_polyphemus_(horseshoe_crab)", "Loligo_pealeii_(longfin_squid)", "Lophius_americanus_(goosefish)", "Macrouridae", "Malacoraja_senta_(smooth_skate)", "Maurolicus_weitzmani_(Weitzmans_pearlsides)", "Melanogrammus_aeglefinus_(haddock)", "Melanostigma_atlanticum_(Atlantic_soft_pout)", "Menidia_menidia_(Atlantic_silverside)", "Menticirrhus_americanus_(southern_kingfish)")


specnames_2 <- c("Menticirrhus_saxatilis_(northern_kingfish)", "Merluccius_albidus_(offshore_hake)", "Merluccius_bilinearis_(silver_hake)", "Micropogonias_undulatus_(Atlantic_croaker)", "Morone_saxatilis_(striped_bass)", "Mustelus_canis_(smooth_dogfish)", "Myctophidae", "Myliobatis_freminvillii_(bullnose_ray)", "Myoxocephalus_octodecemspinosus_(longhorn_sculpin)", "Opisthonema_oglinum_(Atlantic_thread_herring)", "Orthopristis_chrysoptera_(pigfish)", "Ovalipes_ocellatus_(lady_crab)", "Ovalipes_stephensoni_(coarsehand_lady_crab)", "Pandalus_borealis_(northern_shrimp)", "Pandalus_montagui_(Aesop_shrimp)", "Paralichthys_dentatus_(summer_flounder)", "Pasiphaea_multidentata_(pink_glass_shrimp)", "Penaeus", "Peprilus_triacanthus_(butterfish)", "Phycis_chesteri_(longfin_hake)", "Placopecten_magellanicus_(sea_scallop)", "Pollachius_virens_(pollock)", "Polymixia_lowei_(beardfish)", "Pomatomus_saltatrix_(bluefish)", "Prionotus_carolinus_(northern_searobin)", "Prionotus_evolans_(striped_searobin)", "Pseudopleuronectes_americanus_(winter_flounder)", "Raja_eglanteria_(clearnose_skate)", "Rhizoprionodon_terraenovae_(Atlantic_sharpnose_shark)", "Sardinella_aurita_(Spanish_sardine)", "Scomber_scombrus_(Atlantic_mackerel)", "Scophthalmus_aquosus_(windowpane)", "Sebastes_fasciatus_(Acadian_redfish)", "Selene_setapinnis_(Atlantic_moonfish)", "Sphoeroides_maculatus_(northern_puffer)", "Squalus_acanthias_(spiny_dogfish)", "Stenotomus_chrysops_(scup)", "Stephanolepis_hispida_(planehead_filefish)", "Syngnathus_fuscus_(northern_pipefish)", "Synodus_foetens_(inshore_lizardfish)", "Tautogolabrus_adspersus_(cunner)", "Trachinocephalus_myops_(snakefish)", "Trachurus_lathami_(rough_scad)", "Trichiurus_lepturus_(Atlantic_cutlassfish)", "Triglops_murrayi_(moustache_sculpin)", "Urophycis_chuss_(red_hake)", "Urophycis_regia_(spotted_hake)", "Urophycis_tenuis_(white_hake)", "Zenopsis_conchifera_(buckler_dory)", "Zoarces_americanus_(ocean_pout)")

specnames <- c(specnames_1, specnames_2)
```


```{r weighted centroid}
#first time chunk 

spec_coords_1 <- data.frame(name=rep(0,108), x=rep(0,108), y=rep(0,108))
for (i in 1:108) { #for each species in spec_list
  wt.try <- wt.centroid(total_sp_1, specnames[i])
  name <- specnames[i]
  xcord <- wt.try@coords[1]
  ycord <- wt.try@coords[2] 
  spec_coords_1[i,] = c(name, xcord, ycord)
}

#For the 2nd time chunk

spec_coords_2 <- data.frame(name=rep(0,108), x=rep(0,108), y=rep(0,108))
for (i in 1:108) { #for each species in spec_list
  wt.try <- wt.centroid(total_sp_2, specnames[i])
  name <- specnames[i]
  xcord <- wt.try@coords[1]
  ycord <- wt.try@coords[2] 
  spec_coords_2[i,] = c(name, xcord, ycord)
}
```

```{r}
all_specs <- cbind(spec_coords_1,spec_coords_3=2)

colnames(spec_coords_1) <- c("name", "Longitude", "Latitude")
colnames(spec_coords_2) <- c("name", "Longitude", "Latitude")

spec_coords_1$Latitude <- as.numeric(spec_coords_1$Latitude)
spec_coords_1$Longitude <- as.numeric(spec_coords_1$Longitude)

spec_coords_2$Latitude <- as.numeric(spec_coords_2$Latitude)
spec_coords_2$Longitude <- as.numeric(spec_coords_2$Longitude)

```

this calculates the geodesic distance between the two points

```{r geodesic distance}

dist_total <- as.data.frame(distGeo(spec_coords_1[2:3], spec_coords_2[2:3]))
dist_total <- cbind(specnames, dist_total)
dist_total <- cbind(dist_total, spec_coords_1)
dist_total <- cbind(dist_total, spec_coords_2)

write.csv(dist_total, "dist_total.csv")

#lets save all of the mean centroid coordinates to map later
write.csv(spec_coords_1, "spec_coords_1.csv")
write.csv(spec_coords_2, "spec_coords_2.csv")
```






#"ATLANTOPANDALUS.PROPINQVUS","CANCER.PLEBEJUS", "CHACEON.QUINQUEDENS",
#need to fix this because some things are spelled incorrectly etc 
#specnames from Batt 2017
# `Atlantopandalus propinqvus`, `Hippoglossoides platessoides`,`Hippoglossina oblonga`, `Leptoclinus maculatus`, `Lopholatilus chamaeleonticeps`, `Lumpenus lampretaeformis`, `Lycenchelys verrillii`, `Macrorhamphosus scolopax`, 
ydata <- ydata %>% 
   select(`Acipenser oxyrinchus`, `Alosa aestivalis`, `Alosa pseudoharengus`, `Alosa sapidissima`, `Amblyraja radiata`, `Ammodytes dubius`, `Anarhichas lupus`, `Anchoa hepsetus`, `Anchoa mitchilli`, `Antigonia capros`, `Argentina silus`, `Argentina striata`, `Aspidophoroides monopterygius`,`Bairdiella chrysoura`, `Bathynectes longispina`, `Bathypolypus arcticus`, `Brevoortia tyrannus`, `Brosme brosme`, `Cancer borealis`, `Cancer plebejus`, `Carcharhinus plumbeus`, `Centropristis philadelphica`, `Centropristis striata`, `Chaceon quinquedens`, `Chionoecetes opilio`, `Chlorophthalmus agassizi`, `Citharichthys arctifrons`, `Clupea harengus`, `Conger oceanicus`, `Crangon septemspinosa`, `Cryptacanthodes maculatus`, `Cyclopterus lumpus`, `Cynoscion regalis`, `Dasyatis centroura`, `Dasyatis say`, `Decapterus punctatus`, `Dichelopandalus leptocerus`, `Dipturus laevis`, `Enchelyopus cimbrius`, `Etropus microstomus`, `Etrumeus teres(sadina)`, `Foetorepus agassizii`, `Gadus morhua`, `Glyptocephalus cynoglossus`, `Gymnura altavela`, `Helicolenus dactylopterus`, `Hemitripterus americanus`, `Hippoglossus hippoglossus`, `Homarus americanus`, `Illex illecebrosus`, `Lagodon rhomboides`, `Larimus fasciatus`, `Lebbeus polaris`, `Leiostomus xanthurus`, `Lepophidium profundorum`, `Leucoraja erinacea`, `Leucoraja garmani`, `Leucoraja ocellata`, `Limanda ferruginea`, `Limulus polyphemus`, `Liparis atlanticus`, `Lithodes maja`, `Loligo pealeii`, `Lolliguncula brevis`, `Lophius americanus`, `Malacoraja senta`, `Maurolicus weitzmani`, `Melanogrammus aeglefinus`, `Melanostigma atlanticum`, `Menidia menidia`, `Menticirrhus americanus`, `Menticirrhus saxatilis`, `Merluccius albidus`, `Merluccius bilinearis`, `Micropogonias undulatus`, `Monolene sessilicauda`, `Morone saxatilis`, `Mustelus canis`, `Myliobatis freminvillii`, `Myoxocephalus aenaeus`, `Myoxocephalus octodecemspinosus`, `Myxine glutinosa`, `Nemichthys scolopaceus`, `Ophidion marginatum`, `Osmerus mordax`, `Ovalipes ocellatus`, `Ovalipes stephensoni`, `Pandalus borealis`, `Pandalus montagui`, `Paralichthys dentatus`, `Parasudis truculenta`, `Pasiphaea multidentata`, `Peprilus triacanthus`, `Peristedion miniatum`, `Petromyzon marinus`, `Phycis chesteri`, `Placopecten magellanicus`, `Pollachius virens`, `Polymixia lowei`, `Pomatomus saltatrix`, `Pontinus longispinis`, `Pontophilus norvegicus`, `Prionotus alatus`, `Prionotus carolinus`, `Prionotus evolans`, `Pseudopleuronectes americanus`, `Raja eglanteria`, `Scomber scombrus`, `Scophthalmus aquosus`, `Scyliorhinus retifer`, `Sebastes fasciatus`, `Sicyonia brevirostris`, `Squalus acanthias`, `Squatina dumeril`, `Stenotomus chrysops`, `Syacium papillosum`, `Symphurus plagiusa`, `Synagrops bellus`, `Syngnathus fuscus`, `Synodus foetens`, `Synodus poeyi`, `Tautogolabrus adspersus`, `Trachinocephalus myops`, `Trachurus lathami`, `Trichiurus lepturus`, `Triglops murrayi`, `Trinectes maculatus`, `Ulvaria subbifurcata`, `Urophycis chuss`, `Urophycis regia`, `Urophycis tenuis`, `Zenopsis conchifera`, `Zoarces americanus`)


specNames <- c("ACIPENSER.OXYRYNCHUS"`, `"ALOSA.AESTIVALIS", "ALOSA.PSEUDOHARENGUS", "ALOSA.SAPIDISSIMA", "AMBLYRAJA.RADIATA", "AMMODYTES.DUBIUS", "ANARHICHAS.LUPUS", "ANCHOA.HEPSETUS", "ANCHOA.MITCHILLI", "ANTIGONIA.CAPROS", "ARGENTINA.SILUS", "ARGENTINA.STRIATA", "ASPIDOPHOROIDES.MONOPTERYGIUS", "BAIRDIELLA.CHRYSOURA", "BATHYNECTES.LONGISPINA", "BATHYPOLYPUS.ARCTICUS", "BREVOORTIA.TYRANNUS", "BROSME.BROSME", "CANCER.BOREALIS", "CARCHARHINUS.PLUMBEUS", "CENTROPRISTIS.PHILADELPHICA", "CENTROPRISTIS.STRIATA", "CHIONOECETES.OPILIO", "CHLOROPHTHALMUS.AGASSIZI", "CITHARICHTHYS.ARCTIFRONS", "CLUPEA.HARENGUS", "CONGER.OCEANICUS", "CRANGON.SEPTEMSPINOSA", "CRYPTACANTHODES.MACULATUS", "CYCLOPTERUS.LUMPUS", "CYNOSCION.REGALIS", "DASYATIS.CENTROURA", "DASYATIS.SAY", "DECAPTERUS.PUNCTATUS", "DICHELOPANDALUS.LEPTOCERUS", "DIPTURUS.LAEVIS", "ENCHELYOPUS.CIMBRIUS", "ETROPUS.MICROSTOMUS", "ETRUMEUS.TERES", "FOETOREPUS.AGASSIZII", "GADUS.MORHUA", "GLYPTOCEPHALUS.CYNOGLOSSUS", "GYMNURA.ALTAVELA", "HELICOLENUS.DACTYLOPTERUS", "HEMITRIPTERUS.AMERICANUS", "HIPPOGLOSSINA.OBLONGA", "HIPPOGLOSSOIDES.PLATESSOIDES", "HIPPOGLOSSUS.HIPPOGLOSSUS", "HOMARUS.AMERICANUS", "ILLEX.ILLECEBROSUS", "LAGODON.RHOMBOIDES", "LARIMUS.FASCIATUS", "LEBBEUS.POLARIS", "LEIOSTOMUS.XANTHURUS", "LEPOPHIDIUM.PROFUNDORUM", "LEPTOCLINUS.MACULATUS", "LEUCORAJA.ERINACEA", "LEUCORAJA.GARMANI", "LEUCORAJA.OCELLATA", "LIMANDA.FERRUGINEA", "LIMULUS.POLYPHEMUS", "LIPARIS.ATLANTICUS", "LITHODES.MAJA", "LOLIGO.PEALEII", "LOLLIGUNCULA.BREVIS", "LOPHIUS.AMERICANUS", "LOPHOLATILUS.CHAMAELEONTICEPS", "LUMPENUS.LAMPRETAEFORMIS", "LYCENCHELYS.VERRILLII", "MACRORHAMPHOSUS.SCOLOPAX", "MALACORAJA.SENTA", "MAUROLICUS.WEITZMANI", "MELANOGRAMMUS.AEGLEFINUS", "MELANOSTIGMA.ATLANTICUM", "MENIDIA.MENIDIA", "MENTICIRRHUS.AMERICANUS", "MENTICIRRHUS.SAXATILIS", "MERLUCCIUS.ALBIDUS", "MERLUCCIUS.BILINEARIS", "MICROPOGONIAS.UNDULATUS", "MONOLENE.SESSILICAUDA", "MORONE.SAXATILIS", "MUSTELUS.CANIS", "MYLIOBATIS.FREMINVILLII", "MYOXOCEPHALUS.AENAEUS", "MYOXOCEPHALUS.OCTODECEMSPINOSUS", "MYXINE.GLUTINOSA", "NEMICHTHYS.SCOLOPACEUS", "OPHIDION.MARGINATUM", "OSMERUS.MORDAX", "OVALIPES.OCELLATUS", "OVALIPES.STEPHENSONI", "PANDALUS.BOREALIS", "PANDALUS.MONTAGUI", "PARALICHTHYS.DENTATUS", "PARASUDIS.TRUCULENTA", "PASIPHAEA.MULTIDENTATA", "PEPRILUS.TRIACANTHUS", "PERISTEDION.MINIATUM", "PETROMYZON.MARINUS")

specNames <- c("PHYCIS.CHESTERI", "PLACOPECTEN.MAGELLANICUS", "POLLACHIUS.VIRENS", "POLYMIXIA.LOWEI", "POMATOMUS.SALTATRIX", "PONTINUS.LONGISPINIS", "PONTOPHILUS.NORVEGICUS", "PRIONOTUS.ALATUS", "PRIONOTUS.CAROLINUS", "PRIONOTUS.EVOLANS", "PSEUDOPLEURONECTES.AMERICANUS", "RAJA.EGLANTERIA", "SCOMBER.SCOMBRUS", "SCOPHTHALMUS.AQUOSUS", "SCYLIORHINUS.RETIFER", "SEBASTES.FASCIATUS", "SICYONIA.BREVIROSTRIS", "SQUALUS.ACANTHIAS", "SQUATINA.DUMERIL", "STENOTOMUS.CHRYSOPS", "SYACIUM.PAPILLOSUM", "SYMPHURUS.PLAGIUSA", "SYNAGROPS.BELLUS", "SYNGNATHUS.FUSCUS", "SYNODUS.FOETENS", "SYNODUS.POEYI", "TAUTOGOLABRUS.ADSPERSUS", "TRACHINOCEPHALUS.MYOPS", "TRACHURUS.LATHAMI", "TRICHIURUS.LEPTURUS", "TRIGLOPS.MURRATRINECTES.MACULATUS", "ULVARIA.SUBBIFURCATA", "UROPHYCIS.CHUSS", "UROPHYCIS.REGIA", "UROPHYCIS.TENUIS", "ZENOPSIS.CONCHIFERA", "ZOARCES.AMERICANUS")

specNames <- c(spec1, spec2)

xnames <- c("year", "depth", "btemp", "ssalin", "AMO_unsm_month", "NAO_month", "NAO_JFM", "BATHYMETRY", "SBF_GRP", "SEDIMENT")

ydata <- bioandcatch[,specNames]
xdata <- bioandcatch[,xnames]

```

