---
title: "Cleaned_code"
author: "Sarah Roberts"
date: "12/19/2019"
output: html_document
---
##packages
This is to get at if the environment is changing or if species are changing with the environment. 
```{r load pacakges}
#data manip 
library(tidyr)
library(dplyr)
library(naniar) #replace with NA

#stats 
library(mgcv)

#spatial 
library(sp)
library(spatialEco)
library(geosphere)
library(adehabitatHR)
library(rgdal)


#plotting
library(reshape2)
library(ggpubr)
library(ggsci)
library(ggthemes)

```

#standardizing ggplot 
```{r}
ggtheme_Plot <- function() {
  theme(
    plot.title = element_text(size = rel(1), hjust = 0, face = "bold"),
    panel.background = element_blank(),
    strip.background = element_blank(),
    #strip.text       = element_text(size = base_size, face = "italic"),
    panel.border     = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.ticks = element_blank(),
 axis.line = element_line(size = 0.5, linetype = "solid", colour = "black"),
  axis.text.x = element_text(size = 18,
                               angle = 0,
                               face = "plain"),
    axis.text.y = element_text(size = 18),
    axis.title = element_text(size = 18),
    legend.key = element_rect(colour = NA, fill = NA),
    legend.position  = "left",
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 18),
    strip.text.x = element_text(size = 18),
    strip.text = element_text(size = 18) 
 
  )
}

```

#Load data 
```{r load data}

#load data 
fall_catch <- read.csv("~/Documents/ClimateOceanPlanning/Datasets/NEFSC_Bottom_Trawl/22560_FSCSTables_Fall/22560_UNION_FSCS_SVCAT.csv")
fall_x <- read.csv("~/Documents/ClimateOceanPlanning/Datasets/NEFSC_Bottom_Trawl/22560_FSCSTables_Fall/22560_UNION_FSCS_SVSTA.csv")

spring_catch <- read.csv("~/Documents/ClimateOceanPlanning/Datasets/NEFSC_Bottom_Trawl/22561_FSCSTables_spring/22561_UNION_FSCS_SVCAT.csv")
spring_x <- read.csv("~/Documents/ClimateOceanPlanning/Datasets/NEFSC_Bottom_Trawl/22561_FSCSTables_spring/22561_UNION_FSCS_SVSTA.csv")

```

```{r data manip}
#manipulated the catch data to be in tidy format 
fall_catch <- fall_catch %>% 
 group_by_at(vars(-expcatchnum)) %>% # group by everything other than the value column. 
 mutate(row_id=1:n()) %>% ungroup() %>% # build group index
 spread(key=scientific_name, value=expcatchnum)  # spread

fall_catch$id <- as.numeric(as.character(fall_catch$id))

fall_catch_select <- fall_catch %>% 
  dplyr::select(-catch_comment) %>% 
  mutate_if(is.factor, as.integer) %>% 
  group_by(id) %>% 
  replace(is.na(.), 0) %>%
 summarise_each(funs(max)) 

fall_catch_select <- fall_catch_select %>% 
  dplyr::select(-status_code, -svspp, -catchsex, -expcatchwt, -row_id)

#joining the tables 
fall_x$id <- as.numeric(as.character(fall_x$id))
fall <- left_join(fall_catch_select, fall_x, by="id")
fall$season <- "fall"

#doing the same for spring 
spring_catch <- spring_catch %>% 
 group_by_at(vars(-expcatchnum)) %>% # group by everything other than the value column. 
 mutate(row_id=1:n()) %>% ungroup() %>% # build group index
 spread(key=scientific_name, value=expcatchnum)  # spread

spring_catch$id <- as.numeric(as.character(spring_catch$id))

spring_catch_select <- spring_catch %>% 
  dplyr::select(-catch_comment) %>% 
  mutate_if(is.factor, as.integer) %>% 
  group_by(id) %>% 
  replace(is.na(.), 0) %>%
 summarise_each(funs(max)) 

spring_catch_select <- spring_catch_select %>% 
  dplyr::select(-status_code, -svspp, -catchsex, -expcatchwt, -row_id)

#joining the tables 
spring_x$id <- as.numeric(as.character(spring_x$id))
spring <- left_join(spring_catch_select, spring_x, by="id")
spring$season <- "spring"

#combining spring and fall 
#first lets select only the overlapping species/columns 
cols <- intersect(colnames(spring), colnames(fall))
total <- rbind(spring[,cols], fall[,cols])

#now need to filter for the data after 1984 (this is when Batt uses it. There was a change in otter doors to allow more pelagics, and more consistent sampling of southern stratum) http://www.ices.dk/sites/pub/CM%20Doccuments/1997/Y/1997_Y33.pdf

total <- total %>% 
  filter(est_year>1984)
total$myid <- 1:22950
#the original id was wacky
total_full <- total
write.csv(total, file="total.csv")

total_latlondate <- total %>% 
  dplyr::select(myid, decdeg_beglat, decdeg_beglon, est_year, est_month)

write.csv(total_latlondate, file="latlondate_total.csv")
```
Now I need to go back and make a few files. The first one is the NAO/AMO file which is located in the longlatdat_total xl file. Use the index function to get the NAO/AMO values by year. 

The next step is to go in and copy the long lat file to arcmap and use that and the benthic habitat poly 2 shapefile along with the IDENTITY tool to get all of the information into one. 


```{r Adding external data}

#adding AMO/NAO and habitat information 
climate_vars <- read.csv("~/Documents/ClimateOceanPlanning/R/NEFSC_GJAMM_GIT/NAO_AMO_dat.csv")
#lets just get out the important variables (if we get them all then it adds a bunch of unneeded)
climate_vars <- climate_vars %>% 
  dplyr::select(-decdeg_beglat, -decdeg_beglon, -est_year, -est_month)
hab_vars <- read.csv("~/Documents/ClimateOceanPlanning/R/NEFSC_GJAMM_GIT/habitat_dat.csv")

hab_vars <- hab_vars %>% 
  dplyr::select(-decdeg_beg, -decdeg_b_1, -est_year, -est_month)

total <- read.csv("total.csv")

#remove id's where NA 
total <- total %>% 
  filter(!is.na(id)) 

total <- total %>% 
  filter(!is.na(beglat))
#write.csv(total, file="total.csv")
#total <- read.csv("total.csv")
#idk why there was a spatial NA but got rid of it in the other files and here #note this is really frustrating and can't figure out that the id columns are the same so I might just do it in excel 
#note this id is missing spatial points 
#201204013400080000.00


#bind them all together 
total <- left_join(total, climate_vars, by="myid")

total <- left_join(total, hab_vars, by="myid")

#getting rid of some overlapping columns 
total <- total[, !duplicated(colnames(total))]

#replacing spaces with _ in column names 
names(total) <- gsub(" ", "_", names(total))

#trimming dates 
total <- total %>% 
  filter(est_year>1984)

#trimming so just the columns we want 

xdata <- total %>% 
 dplyr::select(myid, est_month, est_day, est_year,surftemp, surfsalin, bottemp,
     botsalin, AMO_unsm_year, NAO_year, NAO_JFM, avgdepth,
     LPI, SLOPE, SBF_GRP, SBFDEPTH, SEDIMENT, SBFDESED, BENHAB, decdeg_beglat, decdeg_beglon, gebco_ra_1, gebco_SD_2)

#too many to do like this for ydata so will just select the columns 
ydata <- total[,8:514]

total <- cbind(ydata, xdata)


#lets check to make sure there aren't any NA's

#saving these files in case we mess up somewhere and need them again 
write.csv(total, file="total.csv")
write.csv(xdata, file="xdata.csv")
write.csv(ydata, file="ydata.csv")


```

Fixing NAs and unidentified fish. Also take some time to make sure x variables have info in most of them  
```{r}

total <- read.csv("total.csv")
ydata <- read.csv("ydata.csv")
xdata <- read.csv("xdata.csv")

total[total == -9999.000] <- NA
xdata[xdata == -9999.000] <- NA

ydata <- ydata %>% 
  dplyr::select(-X)

xdata <- xdata %>% 
  dplyr::select(-X)

xdata <- xdata %>% 
  dplyr::rename(est_month = est_month.x, est_year = est_year.y)

total <- cbind(ydata, xdata)

write.csv(total, file="total.csv")
write.csv(xdata, file="xdata.csv")
write.csv(ydata, file="ydata.csv")

```


#checkpoint 
do we have adataset by myID with all of the x and y data as well as a dataset with just x data and just y data? 
#Full GAM to select species 
#selecting based on model performance 
```{r}
#finding all the species

#manipulated the catch data to be in tidy format 
fall_catch$id <- as.numeric(as.character(fall_catch$id))

#joining the tables 
fall_x$id <- as.numeric(as.character(fall_x$id))

fall_test <- left_join(fall_catch, fall_x, by="id")
fall_test <- fall_test %>% dplyr:::select(id, expcatchnum, scientific_name, est_year, est_month )


spring_catch$id <- as.numeric(as.character(spring_catch$id))

#joining the tables 
spring_x$id <- as.numeric(as.character(spring_x$id))

spring_test <- left_join(spring_catch, spring_x, by="id")
spring_test <- spring_test %>% dplyr:::select(id, expcatchnum, scientific_name, est_year, est_month )

tot_test <- rbind(spring_test, fall_test)

#which species are present in multiple months in both time periods
tot_by_m <- tot_test %>% group_by(scientific_name) %>% 
  summarise(
  months = length(unique(est_month)))


tot_by_m$scientific_name <- gsub(" ", ".", tot_by_m$scientific_name)
tot_by_m$scientific_name <- gsub("\\(", ".", tot_by_m$scientific_name)
tot_by_m$scientific_name <- gsub("\\)", ".", tot_by_m$scientific_name)
tot_by_m$scientific_name <- gsub("\\-", ".", tot_by_m$scientific_name)

write.csv(tot_by_m, file="total_months.csv")


#running full gams with all the species
specnames <- as.vector(tot_by_m$scientific_name)
tot_gam_r_2 <- data.frame(matrix(ncol=3, nrow=1))
colnames(tot_gam_r_2) <- c("names", "rsq", "dev")

for (i in 1:length(specnames)) {     
  tryCatch({#for each species in spec_list
  fit <- gam(as.formula(paste(specnames[i],"~ avgdepth + bottemp + botsalin + SED_TYPE + gebco_SD_2 + surfsalin + surftemp")), data=total, family=nb(link="log"))
  r <- summary(fit)$r.sq
  dev <- summary(fit)$dev.expl
        dev_table <- list("null", r, dev)
        dev_table <- as.data.frame(dev_table)
        colnames(dev_table) <- c("names", "rsq", "dev")
        dev_table$names <- specnames[i]
        
        tot_gam_r_2 <- rbind(tot_gam_r_2, dev_table)
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
        # get coefficents of fit
        #summary_model <- summary(fit)
        #p <- summary_model$p.table
        
        
        #cfit <- fit$coefficients
        #cfit <- as.data.frame(cfit)
        #cfit$names <- paste(specnames[i], rownames(cfit), sep="+")
        #cfit$AIC <- AIC(fit)
        #cfit$Deviance <- dev_table
        #cfit$p_value <- p[,4]
  
       
       # tot_gam_results <- rbind(tot_gam_results, cfit)
        #tot_gam_stand <- rbind(tot_gam_stand, sfit)
} 

#deviance explained might be better than r - squared for non gaussian families
  #https://m-clark.github.io/generalized-additive-models/technical.html#a-detailed-example

#select species whose deviance explained is greater than 50% 
gam_50 <- subset(tot_gam_r, tot_gam_r$dev>=.50)
specnames_r <- as.vector(gam_50$names)

#select species whose deviance explained is greater than 30% 

gam_30 <- subset(tot_gam_r_2, tot_gam_r_2$dev>=.30)
specnames_more <- as.vector(gam_30$names)
```

#individual GAMs
```{r}

#finding deviance explained by each varaible for species with total deviance explained over 30%

tot_gam_dev_50 <- data.frame(matrix(ncol=12, nrow=1))
tot_gam_dev_30 <- data.frame(matrix(ncol=19, nrow=1))

colnames(tot_gam_dev_30) <- c("names", "dev.depth", "dev.bt", "dev.sal", "dev.AMO", "dev.NAO", "dev.sed", "dev.geb", "dev.sst", "dev.ssal", "dev.null", "coef.depth", "coef.bt", "coef.sal", "coef.AMO", "coef.NAO", "coef.geb", "coef.sst", "coef.ssal")
specnames <- specnames_more
for (i in 1:length(specnames)) { #for each species in spec_list
  #fit <- gam(as.formula(paste(specnames[i],"~ avgdepth + bottemp + botsalin + AMO_unsm_year + NAO_year + SED_TYPE + SD3x3shortsal + SD3x3shortbt + SD3x3shortsst + gebco_SD_2")), data=total)
#standardized <- lm.beta(fit)
    tryCatch({
        fit_depth <- gam(as.formula(paste(specnames[i],"~ avgdepth")), data=total, family=nb(link="log"))
          fit_bt <- gam(as.formula(paste(specnames[i],"~ bottemp")), data=total, family=nb(link="log"))
           fit_sal <- gam(as.formula(paste(specnames[i],"~ botsalin")), data=total, family=nb(link="log"))
           fit_AMO <- gam(as.formula(paste(specnames[i],"~ AMO_unsm_year")), data=total, family=nb(link="log"))
           fit_NAO <- gam(as.formula(paste(specnames[i],"~ NAO_year")), data=total, family=nb(link="log"))   
         fit_sed <- gam(as.formula(paste(specnames[i],"~ SEDIMENT")), data=total, family=nb(link="log"))
        fit_geb <- gam(as.formula(paste(specnames[i],"~ gebco_SD_2")), data=total, family=nb(link="log"))
        fit_sst <- gam(as.formula(paste(specnames[i],"~ surftemp")), data=total, family=nb(link="log"))
        fit_ssal <- gam(as.formula(paste(specnames[i],"~ surfsalin")), data=total, family=nb(link="log"))

        fit_null <- gam(as.formula(paste(specnames[i],"~ 1")), data=total, family=nb(link="log"))
        dev.depth <- summary(fit_depth)$dev.expl
        dev.bt <- summary(fit_bt)$dev.expl
        dev.sal <- summary(fit_sal)$dev.expl
        dev.AMO <- summary(fit_AMO)$dev.expl
        dev.NAO <- summary(fit_NAO)$dev.expl
        dev.sed <- summary(fit_sed)$dev.expl
        dev.geb <- summary(fit_geb)$dev.expl
        dev.null <- summary(fit_null)$dev.expl
        dev.sst <- summary(fit_sst)$dev.expl
       dev.ssal <- summary(fit_ssal)$dev.expl

        coef.depth <- summary(fit_depth)$p.coeff[2]
        coef.bt<- summary(fit_bt)$p.coeff[2]
        coef.sal <- summary(fit_sal)$p.coeff[2]
        coef.AMO <- summary(fit_AMO)$p.coeff[2]
        coef.NAO <- summary(fit_NAO)$p.coeff[2]
        coef.geb <- summary(fit_geb)$p.coeff[2]
        coef.sst <- summary(fit_sst)$p.coeff[2]
        coef.ssal <- summary(fit_ssal)$p.coeff[2]
        
        dev_table <- list("null", dev.depth, dev.bt, dev.sal, dev.AMO, dev.NAO, dev.sed, dev.geb, dev.sst, dev.ssal, dev.null, coef.depth, coef.bt, coef.sal, coef.AMO, coef.NAO, coef.sdsal, coef.sdbt, coef.sdsst, coef.geb, coef.sst, coef.ssal)
        dev_table <- as.data.frame(dev_table)
        colnames(dev_table) <- c("names", "dev.depth", "dev.bt", "dev.sal", "dev.AMO", "dev.NAO", "dev.sed", "dev.geb", "dev.sst", "dev.ssal", "dev.null", "coef.depth", "coef.bt", "coef.sal", "coef.AMO", "coef.NAO", "coef.geb", "coef.sst", "coef.ssal")
        dev_table$names <- specnames[i]
        
        tot_gam_dev_30 <- rbind(tot_gam_dev_30, dev_table)
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
        # get coefficents of fit
        #summary_model <- summary(fit)
        #p <- summary_model$p.table
        
        
        #cfit <- fit$coefficients
        #cfit <- as.data.frame(cfit)
        #cfit$names <- paste(specnames[i], rownames(cfit), sep="+")
        #cfit$AIC <- AIC(fit)
        #cfit$Deviance <- dev_table
        #cfit$p_value <- p[,4]
  
       
       # tot_gam_results <- rbind(tot_gam_results, cfit)
        #tot_gam_stand <- rbind(tot_gam_stand, sfit)
} 

```


#mean centroid shifts 
```{r}
total <- read.csv("total.csv")

total_for_sp <- total
names(total)<-make.names(names(total),unique = TRUE)

#Lets select out the time period chunks we need. 
total_sp_85_90 <- subset(total, est_year < 1991)
total_sp_12_17 <- subset(total, est_year > 2011)


coordinates(total_sp_85_90) <- ~ decdeg_beglon + decdeg_beglat

coordinates(total_sp_12_17) <- ~ decdeg_beglon + decdeg_beglat


#first 5 years 
spec_coords_85_90 <- data.frame(name=rep(0,80), x=rep(0,80), y=rep(0,80))
for (i in 1:length(specnames)) { #for each species in spec_list
  tryCatch({#for each species in spec_list
  wt.try <- wt.centroid(total_sp_85_90, specnames[i])
  name <- specnames[i]
  xcord <- wt.try@coords[1]
  ycord <- wt.try@coords[2] 
  spec_coords_85_90[i,] = c(name, xcord, ycord)
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}
#2nd 5 year chunk

spec_coords_12_17 <- data.frame(name=rep(0,80), x=rep(0,80), y=rep(0,80))
for (i in 1:length(specnames)) { #for each species in spec_list
  tryCatch({#for each species in spec_list
    wt.try <- wt.centroid(total_sp_12_17, specnames[i])
  name <- specnames[i]
  xcord <- wt.try@coords[1]
  ycord <- wt.try@coords[2] 
  spec_coords_12_17[i,] = c(name, xcord, ycord)
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

all_specs_2 <- cbind(spec_coords_85_90,spec_coords_12_17)

colnames(spec_coords_85_90) <- c("name", "Longitude_1", "Latitude_1")
colnames(spec_coords_12_17) <- c("name", "Longitude_2", "Latitude_2")

spec_coords_85_90$Latitude_1 <- as.numeric(spec_coords_85_90$Latitude_1)
spec_coords_85_90$Longitude_1 <- as.numeric(spec_coords_85_90$Longitude_1)

spec_coords_12_17$Latitude_2 <- as.numeric(spec_coords_12_17$Latitude_2)
spec_coords_12_17$Longitude_2 <- as.numeric(spec_coords_12_17$Longitude_2)


#this calculates the geodesic distance between the two points

#for 5 year chunks
dist_total_5yr <- as.data.frame(distGeo(spec_coords_85_90[2:3], spec_coords_12_17[2:3]))
dist_total_5yr <- cbind(specnames, dist_total_5yr)
dist_total_5yr <- cbind(dist_total_5yr, spec_coords_85_90)
dist_total_5yr <- cbind(dist_total_5yr, spec_coords_12_17)

write.csv(dist_total_5yr, "dist_total_5yr_3.csv")

#lets save all of the mean centroid coordinates to map later
#write.csv(spec_coords_85_90, "spec_coords_85_90_2.csv")
#write.csv(spec_coords_12_17, "spec_coords_12_17_2.csv")

dist_total_5yr$dist <- dist_total_5yr$`distGeo(spec_coords_85_90[2:3], spec_coords_12_17[2:3])`
dist_total_5yr <- subset(dist_total_5yr, dist_total_5yr$Longitude_1 != 0)

dist_total_5yr <- subset(dist_total_5yr, dist_total_5yr$Longitude_2 != 0)
write.csv(dist_total_5yr, "dist_total_5yr_3.csv")

```

#convex hull 

```{r}
total_big <- rbind(fall, spring)
total_big_1 <- as.data.frame(total_big)
total_big_1 <- subset(total_big_1, total_big_1$est_year >= 1984)

total_big_1 <- total_big_1 %>% 
  dplyr::select(scientific_name, decdeg_beglon, decdeg_beglat) %>% 
  rename(ID=scientific_name)


#check if any specs have less than 5 relocations using dplyr
total_big_1 <- subset(total_big_1, ID %in% names(which(table(ID) >= 5))) 
total_big_1$ID <- factor(total_big_1$ID)

#make a spatial object
total_big_sp <- total_big_1 %>% drop_na()
coordinates(total_big_sp) <- ~ decdeg_beglon + decdeg_beglat

spec_hull_tot <- mcp(total_big_sp[,1], percent=95)
spec_hull_tot_area <- as.data.frame(spec_hull_tot[])


#now for two time chunks
total_big <- rbind(fall, spring)
total_big_1 <- as.data.frame(total_big)
total_big_1 <- subset(total_big_1, total_big_1$est_year >= 1984)

total_big_first <- subset(total_big_1, est_year < 1991)
total_big_second <- subset(total_big_1, est_year > 2011)

#first
total_big_first <- total_big_first %>% 
  dplyr::select(scientific_name, decdeg_beglon, decdeg_beglat) %>% 
  rename(ID=scientific_name)


#check if any specs have less than 5 relocations using dplyr
total_big_first <- subset(total_big_first, ID %in% names(which(table(ID) >= 5)))
total_big_first$ID <- factor(total_big_first$ID)

#make a spatial object
total_big_first_sp <- total_big_first %>% drop_na()
coordinates(total_big_first_sp) <- ~ decdeg_beglon + decdeg_beglat


spec_hull_first <- mcp(total_big_first_sp[,1], percent=95)
spec_hull_first_area <- as.data.frame(spec_hull_first[])
plot(spec_hull_first)

#second
total_big_second <- total_big_second %>% 
  dplyr::select(scientific_name, decdeg_beglon, decdeg_beglat) %>% 
  rename(ID=scientific_name)


#check if any specs have less than 5 relocations using dplyr
total_big_second <- subset(total_big_second, ID %in% names(which(table(ID) >= 5)))
total_big_second$ID <- factor(total_big_second$ID)

#make a spatial object
total_big_second_sp <- total_big_second %>% drop_na()
coordinates(total_big_second_sp) <- ~ decdeg_beglon + decdeg_beglat


spec_hull_second <- mcp(total_big_second_sp[,1], percent=95)
spec_hull_second_area <- as.data.frame(spec_hull_second[])
plot(spec_hull_first)
plot(spec_hull_second)

write.csv(spec_hull_first_area, file="spec_hull_first5yr_area.csv")
write.csv(spec_hull_second_area, file="spec_hull_second5yr_area.csv")

#write these out to shapefiles
writeOGR(spec_hull_first, layer= "homerange_1_5yr", driver="ESRI Shapefile", dsn="spechull")
writeOGR(spec_hull_second, layer= "homerange_2_5yr", driver="ESRI Shapefile", dsn="spechull")

#go into r arcmap and use the add geometry attributes units to find min and max lat 

```

#combining spatial and gam data 
```{r}
tot_gam_dev_30 <- tot_gam_dev_30[2:377,]

tot_gam_dev_30$strong <- colnames(tot_gam_dev_30[2:13])[apply(tot_gam_dev_30[2:13],1,which.max)]

#info with species type 
specnames_30  <- read.csv("specnames_318.csv")
specnames_30 <- specnames_30[,1:9]
specnames_30 <- specnames_30[1:376,]
specnames_30 <- left_join(tot_gam_r_2, specnames_30, by=c("names"))

specnames_30 <- subset(specnames_30, specnames_30$dev >=.3)

#add distance info 
dist_total <- read.csv("dist_total_full_400.csv")
dist_info <- dist_total[,2:3] 
dist_info$dist_15yr <- dist_info$distGeo.spec_coords_1.2.3...spec_coords_2.2.3..
dist_info_5yr <- dist_total_5yr
dist_info_5yr$dist_5yr <- dist_info_5yr$dist
dist_info_5yr <- dist_info_5yr %>% dplyr::select(specnames, dist_5yr)

tot_gam_dev_30$sum_dev <- rowSums(tot_gam_dev_30[,2:13])

#convex info 
convex_1 <- read.csv("spechull/homerange_1.csv")
convex_2 <- read.csv("spechull/homerange_2.csv")
convex_1 <- convex_1 %>% dplyr::select(id, names, area, EXT_MIN_Y, EXT_MAX_Y)
convex_2 <- convex_2 %>% dplyr::select(id, names, area, EXT_MIN_Y, EXT_MAX_Y)
colnames(convex_1) <- c("id", "names", "area_1", "min_y_1", "max_y_1")
colnames(convex_2) <- c("id", "names", "area_2", "min_y_2", "max_y_2")
convex1 <- merge(convex_2, convex_1, by="names")
convex1$area_diff <- convex1$area_2 - convex1$area_1
convex1$min_diff <- convex1$min_y_2 - convex$min_y_1
convex1$max_diff <- convex1$max_y_2 - convex1$max_y_1

convex_1_5 <- read.csv("homerange_1_5yr.csv")
convex_2_5 <- read.csv("homerange_2_5yr.csv")
convex_1_5 <- convex_1_5 %>% dplyr::select(id, area, EXT_MIN_Y, EXT_MAX_Y)
convex_2_5 <- convex_2_5 %>% dplyr::select(id, area, EXT_MIN_Y, EXT_MAX_Y)
colnames(convex_1_5) <- c("names", "area_1", "min_y_1", "max_y_1")
colnames(convex_2_5) <- c("names", "area_2", "min_y_2", "max_y_2")
convex <- merge(convex_2_5, convex_1_5, by="names")
convex$area_diff_5yr <- convex$area_2 - convex$area_1
convex$min_diff_5yr <- convex$min_y_2 - convex$min_y_1
convex$max_diff_5yr <- convex$max_y_2 - convex$max_y_1

convex <- left_join(convex1, convex, by=c("id.x"="names"))

specinfo <- left_join(specnames_30, dist_info, by=c("names"="specnames_dist"))
specinfo <- left_join(specinfo, dist_info_5yr, by=c("names"="specnames"))
specinfo <- left_join(specinfo, convex, by=c("names"))
#specinfo$dist <- specinfo$`distGeo(spec_coords_1[2:3], spec_coords_2[2:3])`
gam_supp <- left_join(tot_gam_dev_30, specinfo, by=c("names"))

#gam_supp$dist <- gam_supp$distGeo.spec_coords_1.2.3...spec_coords_2.2.3..
gam_supp_orig <- gam_supp
write.csv(gam_supp_orig, "gam_supp.csv")

#adjusting specnames to exclude families and only include fish (can exclude elasmobranchs and only do teleosts too) Could just do phylum chordata which is all fish and sharks and rays but not octopods, mollusks, gastropods, or cephalopods)
gam_supp<- subset(gam_supp, gam_supp$question != "family")
#gam_supp<- subset(gam_supp, gam_supp$question != "yes")
gam_supp<- subset(gam_supp, gam_supp$fish_type == "fish" |  gam_supp$fish_type == "elasmobranch")
#| gam_supp$question != "yes")

#try combining benthic and demersal 
gam_supp$depth_saba_2 <- ifelse(gam_supp$depth_saba == "benthic", "demersal", as.character(gam_supp$depth_saba))
#gam_supp<- subset(gam_supp, gam_supp$depth_saba != "reef")

gam_supp$strong.dev.2 <- ifelse(gam_supp$strong == "dev.sed", "substrate", ifelse(gam_supp$strong == "dev.bt", "bottemp", ifelse(gam_supp$strong == "dev.depth", "depth", ifelse(gam_supp$strong == "dev.AMO" | gam_supp$strong == "dev.NAO", "climate", ifelse(gam_supp$strong == "dev.sal", "salinity", ifelse(gam_supp$strong == "dev.sdsal", "sal variability", ifelse(gam_supp$strong=="dev.sdsst", "sst variability", ifelse(gam_supp$strong=="dev.sst", "sst", ifelse(gam_supp$strong=="dev.ssal", "ssal", ifelse(gam_supp$strong=="dev.geb", "rugosity", NA))))))))))

gam_supp <- gam_supp %>% dplyr::select(-dev.null)
#gam_supp <- subset(gam_supp, gam_supp$names !="Hypleurochilus.geminatus..crested.blenny.")
gam_supp_3 <- gam_supp[1:285,]
gam_supp_orig <- gam_supp
write.csv(gam_supp_5yr, "gam_supp_5yrcurrent.csv")


gam_supp <- gam_supp_orig
gam_supp_5yr <- subset(gam_supp, (!is.na(gam_supp$dist_5yr)))

#without variability 
gam_supp_5yr <- gam_supp_5yr %>%  dplyr::select(-strong, -dev.AMO, -dev.NAO, -dev.sdbt, -dev.sdsal, -dev.sdsst)
gam_supp_5yr$strong <- colnames(gam_supp_5yr[2:8])[apply(gam_supp_5yr[2:8],1,which.max)]

gam_supp_5yr$strong.dev.2 <- ifelse(gam_supp_5yr$strong == "dev.sed", "substrate", ifelse(gam_supp_5yr$strong == "dev.bt", "bottom temperature", ifelse(gam_supp_5yr$strong == "dev.depth", "depth", ifelse(gam_supp_5yr$strong == "dev.sal", "bottom salinity", ifelse(gam_supp_5yr$strong=="dev.sst", "surface temperature", ifelse(gam_supp_5yr$strong=="dev.ssal", "surface salinity", ifelse(gam_supp_5yr$strong=="dev.geb", "rugosity", NA)))))))
```

```{r}
gam_supp_5yr <- read.csv("gam_supp_5yrcurrent.csv")
```

#statistics 
```{r}

pairwise.wilcox.test(as.numeric(gam_supp$dist_5yr), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$dist_15yr), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$area_diff_5yr), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$min_diff_5yr), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$max_diff_5yr), gam_supp$depth_saba_2)


pairwise.wilcox.test(as.numeric(gam_supp$area_diff), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$min_diff), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$max_diff), gam_supp$depth_saba_2)


pairwise.wilcox.test(as.numeric(gam_supp$dist), gam_supp$strong.dev.2)
pairwise.wilcox.test(as.numeric(gam_supp$dev.bt), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$dev.sed), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$dev.depth), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$dev.sal), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$dev.NAO), gam_supp$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp$dev.AMO), gam_supp$depth_saba_2)

pairwise.wilcox.test(as.numeric(gam_supp_5yr$dist), gam_supp_5yr$strong.dev.2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.bt), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.sed), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.depth), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.geb), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.sal), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.NAO), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.AMO), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.sst), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.sdsal), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.sdsst), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.sdbt), gam_supp_5yr$depth_saba_2)
pairwise.wilcox.test(as.numeric(gam_supp_5yr$dev.ssal), gam_supp_5yr$depth_saba_2)


```

#plots

```{r}

cbbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#D55E00", "#0072B2", "#CC79A7")
ggplot(data=gam_supp_5yr, aes(x= strong.dev.2,  group=depth_saba_2)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.5) +
    labs(y = "Percent", x = "strongest variable", fill="strog_2") + 
  coord_flip() +
    facet_wrap(~depth_saba_2) +
    scale_y_continuous(labels = scales::percent) + theme(legend.position = "none")  +
    ggtheme_Plot()  
  ggsave("barplot_gam_more.jpg", plot=last_plot(), width = 20, height=8, units=c("in"), dpi=500)

cbbPalette <- c("#999999", "#E69F00","#E69F00", "#56B4E9", "#009E73", "#D55E00", "#0072B2", "#CC79A7")



gam_supp_2 <- gam_supp_5yr
colnames(gam_supp_2) <- c("names", "depth", "bottom temperature", "bottom salinity", "substrate", 'rugosity', 'surface temperature', 'surface salinity')

dat.m <- melt(gam_supp_2,id.vars='names', measure.vars=c("depth", "substrate", 'surface temperature',"bottom temperature",  "bottom salinity", 'surface salinity'))
levels(dat.m$variable) <- gsub(" ", "\n", levels(dat.m$variable))

dat.m$value2 <- dat.m$value*100
data.m <- left_join(gam_supp_5yr, dat.m, by=c("names"))
data.m <- as.data.frame(data.m)
ggbarplot(data.m, x="variable", y="value2", add="mean", facet.by = "depth_saba_2", fill = "variable", label = "value2", lab.nb.digits=1, color="variable", orientation = "horiz", lab.size = 8, lab.hjust=1, xlim=c(0,100)) +
    labs(y = "Average Deviance explained (%)", x = "Variable")+ 
 theme_classic(base_size=22) +
  theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1)), 
        strip.text = element_text(size = rel(1)), 
        plot.margin = margin(2, 2, 2, 2, "cm")) + scale_fill_simpsons() + scale_color_simpsons() +
  theme(legend.position = "none") +
  ggsave(path = "figs", filename= "bar_deviance_by_spec_var.jpg", plot=last_plot(), width = 20, height=10, units=c("in"), dpi=500)


my_comparisons <- list( c("pelagic", "reef"), c("demersal", "reef"), c("demersal", "pelagic"))



#all deviance explained 
ggplot(data=data.m, aes(x=depth_saba_2, y=value, fill=depth_saba_2)) + geom_boxplot() +
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=3,show_guide = FALSE) +
    labs(y = "Deviance explained (%)", x = "Species type") + stat_compare_means(comparisons = my_comparisons, method="wilcox.test", label = "p.format", size=7) + labs(fill = "Species Type") + 
  facet_wrap(~variable) + 
 theme_classic(base_size=22) +
  theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1)), 
        strip.text = element_text(size = rel(1)), 
        panel.spacing = unit(1, "cm")) + 
  theme(legend.position = "none") + scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  ggsave(path = "figs", filename= "deviance_by_spec_var.jpg", plot=last_plot(), width = 20, height=18, units=c("in"), dpi=500)


#all distance 
gam_supp_5yr$area_diff_5yr_m <- gam_supp_5yr$area_diff_5yr*10000 #was in hectares
gam_supp_5yr$centroid_shift_km <- gam_supp_5yr$dist_5yr/1000 #was in meters
gam_supp_3 <- gam_supp_5yr %>% dplyr::select(names, centroid_shift_km, area_diff_5yr_m, min_diff_5yr, max_diff_5yr)
colnames(gam_supp_3) <- c("names", "Mean centroid shift (km)", "Range size change (meters squared)", "Southern range shifted (latitude)", "Northern range shifted (latitude)")

dat.m.2 <- melt(gam_supp_3,id.vars='names', measure.vars=c("Mean centroid shift (km)", "Range size change (meters squared)", "Southern range shifted (latitude)", "Northern range shifted (latitude)"))

data.m.2 <- left_join(gam_supp_5yr, dat.m.2, by=c("names"))
data.m.2 <- as.data.frame(data.m.2)

ggplot(data=data.m.2, aes(x=depth_saba_2, y=value, fill=depth_saba_2)) + geom_boxplot() +
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=3,show_guide = FALSE) +
    labs(y = "", x = "Species type") + stat_compare_means(comparisons = my_comparisons, method="wilcox.test", label = "p.format", size=7) + labs(fill = "Species Type") + 
  facet_wrap(~variable, scales = "free") + 
 theme_classic(base_size=22) +
  theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1)), 
        strip.text = element_text(size = rel(1)), 
        panel.spacing = unit(1, "cm")) + 
  theme(legend.position = "none") + scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9")) + 
  ggsave(path = "figs", filename= "distance_by_spec_var.jpg", plot=last_plot(), width = 20, height=20, units=c("in"), dpi=500)
```

